---
title: "DATA 607 Extra Credit - Global Baseline Estimate Recommendation System"
author: "Susanna Wong"
date: "2023-04-25"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Prompt

Using the information you collected on movie ratings, implement a Global Baseline Estimate recommendation system in R.  
Most recommender systems use personalized algorithms like “content management” and “item-item collaborative filtering.” Sometimes non-personalized recommenders are also useful or necessary. One of the best non-personalized recommender system algorithms is the “Global Baseline Estimate".
The job here is to use the survey data collected and write the R code that makes a movie recommendation using the Global Baseline Estimate algorithm.  Please see the attached spreadsheet for implementation details.

In this assignment 2, I asked my friends and family to rate 6 movies on a scale of 1 to 5. I took their response and created a table. Below is the link to the Google Form. <https://docs.google.com/forms/d/e/1FAIpQLSd3ht1QlE3axt5LtI3KMkc4skkR9r6wn8PqtPYvVQphLhuQug/viewform>

# Connect to SQL

```{r}
library(tidyverse)
library(DT)
library(RMySQL)
library(DBI)
dbconnect = dbConnect(RMySQL::MySQL(),
                           dbname='Movie Rating',
                           host='localhost',
                           port=3306,
                           user='root',
                           password = rstudioapi::askForPassword("Database password"))

rating <- dbGetQuery(dbconnect, "SELECT * FROM `Movie Rating`.`Movie Rating`")
datatable(rating, options = list(scrollX = TRUE))
```

# Movie 
Below is a table of the movies that each person is rating.

```{r}
movies <- dbGetQuery(dbconnect, "SELECT * FROM `Movie Rating`.Movie;")
datatable(movies, caption = 'Table 1: Movie title and Year', options = list(scrollX = TRUE))
```

# Reviewer Information 
Here is table of first name, age, and gender of each person.

```{r}
person <- dbGetQuery(dbconnect, "SELECT * FROM `Movie Rating`.Person;")
datatable(person, caption = 'Table 2: Information of Reviewers', options = list(scrollX = TRUE))
```

# Calculate the average user rating and average movie rating
Here is a table of each person's rating of each movie. There are missing values for those people who has not seen a movie. 4 people have never seen "Everything Everywhere All at Once" and "Scream". 2 people has never seen "Top Gun: Maverick".

```{r }
rating <- dbGetQuery(dbconnect, "SELECT * FROM `Movie Rating`.`Movie Rating`")
datatable(rating, caption = 'Table 3: Ratings of each movie.', options = list(scrollX = TRUE))
```

Below is a table of the average rating of each movie and average rating for each user. Some people did not see certain movies. So there are missing values. I used "na.rm = TRUE" to deal with missing values. Avengers: Endgame has the highest average rating and Scream has the lowest average rating. Both movies "Everything Everywhere All at Once" and "Scream" had 4 missing values each. Yet, "Everything Everywhere All at Once" had a higher average rating than "Scream" did.

```{r}

rating$user_avg <-apply(rating[,5:10],1,mean,na.rm=TRUE) # Calculate the average user rating

average_movie <- summarize_all(rating, mean, na.rm=TRUE) # Calculate average movie rating in a separate dataframe (Calculate the mean of each column)

average_rating <- rbind(rating, average_movie) # Combine the two dataframes above

average_rating[9,1] <- NA

average_rating[9,"Gender"] <- "average"

datatable(average_rating, options = list(scrollX = TRUE))
```

# Calculate Each Movie's Rating Relative to Average and User's rating relative to average

Movies's rating relative to average = Movie average - Mean Movie

User's rating relative to average = User average - Mean Movie

```{r}
# Calculate Each User's Rating Relative to Average 
average_rating <- average_rating %>%
  mutate(user_relative_rating = user_avg - average_rating[9,11])
  
average_rating <- average_rating %>% 
  add_row(Gender = "movie_relative_rating")

# Calculate Each Movie's Rating Relative to Average

average_rating[10,5] <- average_rating[9,5] - average_rating[9,11]
average_rating[10,6] <- average_rating[9,6] - average_rating[9,11]
average_rating[10,7] <- average_rating[9,7] - average_rating[9,11]
average_rating[10,8] <- average_rating[9,8] - average_rating[9,11]
average_rating[10,9] <- average_rating[9,9] - average_rating[9,11]
average_rating[10,10] <- average_rating[9,10] - average_rating[9,11]
```

# Calculate User's Rating for Movies They did not watch 

There are missing values for those people who has not seen a movie. 4 people have never seen "Everything Everywhere All at Once" and "Scream". 2 people has never seen "Top Gun: Maverick". 

For those missing values, we estimate their rating for those movies using the “Global Baseline Estimate". 

Global Baseline Estimate = Mean Movie Rating + Movies's rating relative to average + User's rating relative to average

```{r}
for(i in 1:ncol(average_rating)){
  if(is.numeric(average_rating[[i]])){
    na <- is.na(average_rating[[i]])
    average_rating[na, i] <- average_rating[9,11] + average_rating[10,i] + average_rating[na, 12]
  }
}

average_rating[,5:12] <-round(average_rating[,5:12],2)

datatable(average_rating, options = list(scrollX = TRUE))
```

# Conclusion

The following people did not watch "Everything Everywhere All at Once": Mel, Ivan, Jack, and Bob. Out of all four people, Ivan has the highest estimate. 

The following people did not watch "Scream": Jenny, James, Ivan, and Jack. Out of all four people, Jenny has the highest estimate. 

Ivan did not watch "Everything Everywhere All at Once" and "Scream". However, he has a higher estimate for the first movie. He is likely to like this movie more. 


# Source

The link below shows me how to use a loop to replace a NA value. The example in the link replaces the NA value with the median of the column. I wanted to replace the NA value with estimated movie rating for a user. 

https://stackoverflow.com/questions/67075477/replacing-na-values-in-a-data-frame-with-the-median-using-a-for-loop
